---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, comment = "", prompt = TRUE)
knitr::opts_chunk$set(out.width="65%", out.height="65%", fig.align="center") 
```

```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(gridExtra)
library(ISLR2)
library(GGally)
library(effectsize)
library(MASS)
library(FNN)
library(dvmisc)
library(boot)
library(caret)
library(corrplot)
library(glmnet)
library(MLmetrics)
library(themis)
library(stargazer)
library(cluster)
library(forcats)
library(glue)
library(corrplot)
library(car)

Sys.setenv('R_MAX_VSIZE'=32000000000)
```

# EDA

## Inputing data

```{r}
df <- read_csv("/Users/jakegwinn/Documents/argonne_all/merged.landmark.csv")
```

```{r}
data <- read_tsv("/Users/jakegwinn/Documents/argonne_all/argonne/data/raw_data/y_data/response.tsv")
```

```{r}
df
```

```{r}
df <- df %>% 
    mutate(
        source = as.factor(source), 
        improve_chem_id = as.factor(improve_chem_id), 
        improve_sample_id = as.factor(improve_sample_id)
    )
```

```{r}
data
```

## Correlation Analysis

```{r}
data 
nums <- unlist(lapply(data, is.numeric), use.names = F)

data_nums <- data[, nums] %>% drop_na()
```

```{r}
d.corr <- cor(data_nums)

corrplot(d.corr)
# d.corr
```

```{r}
d.corr
```

```{r}
nums <- unlist(lapply(df, is.numeric), use.names = F)

df_nums <- df[, nums]
# df_nums
df_nums_small <- df_nums %>% dplyr::select(-contains("mordred")) %>% drop_na()

df_nums_small %>% dplyr::select(aac1, auc1)
```

```{r}
x_data <- df[, 16:973]
df
# x_data[ ,ncol(x_data)]
# ncol(x_data)
# lapply(x_data, is.numeric)
x.corr <- cor(x_data)
x.corr[1,1]
```

```{r}
df_latter <- df[,973:ncol(df)] %>% drop_na()
```

```{r}
latter.corr <- cor(df_latter, use = "complete.obs")
latter.corr
```

```{r}

```

```{r}
significant <- list()

for (i in 1:length(latter.corr[1,])){
    for (j in 1:length(latter.corr[,1])){
        
        row <- rownames(latter.corr)[i]
        col <- colnames(latter.corr)[j]
        # print(abs(latter.corr[i,j]))
        
        if (!is.na(abs(latter.corr[i,j]))) {
            if ((abs(latter.corr[i,j]) > 0.99)){
            # print(x.corr[i,j])
            
            
            
                row <- rownames(latter.corr)[i]
                col <- colnames(latter.corr)[j]  
                
                if (abs(latter.corr[i,j]) == 1) { 
                    if(row!=col) { 
                        rowcol <- c(row, col)            
                        val <- list(c(rowcol, latter.corr[i,j]))
                        significant <- c(significant, val)
                    }
                }
                
                else{ 
                    rowcol <- c(row, col)            
                
                    val <- list(c(rowcol, latter.corr[i,j]))
                # print(val)
                # rowcol <- c(row, col, x.corr[i,j])
                    significant <- c(significant, val)
                }
                
                
            }
        }
    }
}

length(significant)
significant
```

```{r}
df_latter %>% 
    ggplot(aes(x=mordred.AATSC0m, y=mordred.AATSC0Z)) + 
    geom_point() + 
    geom_smooth()
```


```{r}
auc <- df[,"auc"]

model_data <- cbind(x_data, auc)
model_data
```

```{r}
model <- lm(auc ~. ,data = model_data)
```

```{r}
summary(model)$coef
```

```{r}
x.corr
```

```{r}
attributes(alias(model)$Complete)$dimnames[[1]]
```

```{r}
length(x.corr[,1])
```

```{r}
# vif(model)
```

```{r}
significant <- list()

for (i in 1:length(x.corr[1,])){
    for (j in 1:length(x.corr[,1])){
        
        row <- rownames(x.corr)[i]
        col <- colnames(x.corr)[j]
        
        if ((abs(x.corr[i,j]) > 0.8)){
            # print(x.corr[i,j])
            
            
            
            row <- rownames(x.corr)[i]
            col <- colnames(x.corr)[j]  
            
            if (abs(x.corr[i,j]) == 1) { 
                if(row!=col) { 
                    rowcol <- c(row, col)            
                    val <- list(c(rowcol, x.corr[i,j]))
                    significant <- c(significant, val)
                }
            }
            
            else{ 
                rowcol <- c(row, col)            
            
                val <- list(c(rowcol, x.corr[i,j]))
            # print(val)
            # rowcol <- c(row, col, x.corr[i,j])
                significant <- c(significant, val)
            }
            
            
        }
    }
}

length(significant)
significant

```

```{r}
x_data %>% 
    ggplot(aes(x = BUB1B, y = NUSAP1)) + 
    geom_point() + 
    geom_smooth()
```

```{r}
x_data %>% 
    ggplot(aes(x = `IKZF1`, y = `TJP1`)) + 
    geom_point() + 
    geom_smooth()
```

```{r}
df %>% 
    ggplot(aes(x = dss1, y = auc1)) + 
    geom_point(method=lm, level = 0.95, aes(alpha = 0.5, )) + 
    geom_smooth() + 
    labs(
        title = "dss1 vs auc1"
    )
    
```

```{r}
df %>% 
    ggplot(aes(x = dss1, y=auc)) + 
    geom_point(method=lm, level = 0.95, aes(alpha = 0.5, )) +
    geom_smooth()
```

```{r}
df %>% 
    ggplot(aes(x = aac1, y=auc1)) + 
    geom_point(method=lm, level = 0.95, aes(alpha = 0.5, )) +
    geom_smooth()
```

```{r}
df %>% 
    ggplot(aes(x = dss1, y=aac1)) + 
    geom_point(method=lm, level = 0.95, aes(alpha = 0.5, )) +
    geom_smooth()
```

# Drug Analysis

```{r}
# data.cor <- cor(df_nums_small)
```

```{r}
# df[df$improve_chem_id == "Drug_1078",]
```

## EDA

### Overall Stats of Study, drug combos

```{r}
df %>% 
    group_by(source, improve_chem_id) %>% 
    summarize(
        mean_auc = mean(auc),
        auc_var = var(auc),
        count = n()
    ) %>% 
    arrange(improve_chem_id, desc(count) ) 
    
```

### Most common drug and sample combos

```{r}
df %>% 
    group_by(improve_chem_id, improve_sample_id) %>% 
    summarize(
        mean_auc = mean(auc), 
        count = n()
    ) %>% 
    arrange(desc(count))
```

```{r}
df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        mean_auc = mean(auc), 
        count = n()
    ) %>% 
    arrange(desc(count)) %>% 
    ggplot(aes(x = improve_chem_id, y = count)) + 
    geom_bar(stat = "Identity")

```

```{r}
by_study_box <- function(drug) { 
    df[df$improve_chem_id == drug,] %>% 
    ggplot(aes(x = improve_chem_id, y = auc, fill = source)) + 
    geom_boxplot()
}
```

```{r, fig.width=7, fig.height=7}
by_study_box("Drug_988")
```

## Drug_1078 Investigation

### Histograms by study

```{r}
drg = "Drug_1078"
df[df$improve_chem_id == drg,] %>% 
    ggplot(aes(x = auc, fill = source)) + 
    geom_histogram(alpha = 0.6, position = "identity", bins = 35) + 
    facet_wrap(~source) + 
    labs(
        title = glue("{drg} in Different Studies"), 
    ) 

# ggsave("drug988_by_study.png", path = "/Users/jakegwinn/Documents/argonne_all/argonne/plots")
```

### Number of distinct samples per study

```{r}
df[df$improve_chem_id == "Drug_1078",] %>% 
    group_by(source) %>% 
    summarize(
        num_samples = n_distinct(improve_sample_id)
    )
    
```

### Getting samples that are in all studies

```{r}
symdiff <- function( x, y) { setdiff( union(x, y), intersect(x, y))}
```

```{r}
d1078_GDSCv1 <- df[df$improve_chem_id == "Drug_1078" & df$source == "GDSCv1",]
d1078_CTRPv2 <- df[df$improve_chem_id == "Drug_1078" & df$source == "CTRPv2",]
d1078_GDSCv2 <- df[df$improve_chem_id == "Drug_1078" & df$source == "GDSCv2",]
d1078_ccle <- df[df$improve_chem_id == "Drug_1078" & df$source == "CCLE",]


```

```{r}
samp_not_GDSCv2 <- setdiff(d1078_GDSCv1$improve_sample_id, d1078_GDSCv2$improve_sample_id)
length(samp_not_GDSCv2)
```

```{r}
samp_not_CTRPv2 <- setdiff(d1078_GDSCv1$improve_sample_id, d1078_CTRPv2$improve_sample_id)
length(samp_not_CTRPv2)
```

```{r}
setdiff(samp_not_CTRPv2, samp_not_GDSCv2)
```

```{r}
setdiff(samp_not_GDSCv2, samp_not_CTRPv2)
```
#### Samples with weirdly low auc

```{r}
super_low_auc <- d1078_GDSCv1 %>% 
    dplyr::select(source, improve_sample_id, auc) %>% 
    arrange((auc)) 
# %>% 
#     dplyr::select(improve_sample_id)

super_low_auc
```

```{r}
df[df$improve_chem_id == "Drug_1078" & df$source == "GDSCv1" & df$improve_sample_id %in% super_low_auc$improve_sample_id, ]
```

## All super low auc samples are in other studies that do not measure them as low

```{r}
d1078_CTRPv2[d1078_CTRPv2$improve_sample_id %in% super_low_auc$improve_sample_id,]
d1078_GDSCv2[d1078_GDSCv2$improve_sample_id %in% super_low_auc$improve_sample_id,]
```

```{r}
d1078_GDSCv1[d1078_GDSCv1$improve_sample_id %in% super_low_auc$improve_sample_id,]
```

## Looking at 1078 same samples auc distribution

```{r}
same_samps_1078 <- intersect(d1078_ccle$improve_sample_id, intersect(d1078_GDSCv1$improve_sample_id, intersect(d1078_CTRPv2$improve_sample_id, d1078_GDSCv2$improve_sample_id)))
length(same_samps)
```

```{r}
drg = "Drug_1078"
data_1078 <- df[df$improve_chem_id == "Drug_1078" & df$improve_sample_id %in% same_samps_1078,]

data %>% 
    ggplot(aes(x = auc, fill = source)) + 
    geom_histogram(alpha = 0.6, position = "identity", bins = 35) + 
    facet_wrap(~source) + 
    labs(
        title = glue("{drg} in Different Studies with Same Samples") 
    )

# ggsave("drug_1078_by_study_samesamps.png", path = "/Users/jakegwinn/Documents/argonne_all/argonne/plots", dpi = 700)
```
## 1078 super low auc 
```{r}
small_same_samps <- data[data$source == "GDSCv1",] %>% 
    dplyr::select(improve_sample_id, auc) %>% 
    arrange(auc) %>% 
    filter(auc < 0.01)
```

```{r}
data_1078[data$improve_sample_id %in% small_same_samps$improve_sample_id,] %>% 
    dplyr::select(source, improve_sample_id, auc) %>% 
    arrange(improve_sample_id, desc(auc)) 
# %>% 
#     ggplot(aes(x = auc, color = improve_sample_id)) + 
#     geom_boxplot()
```

```{r}
drg = "Drug_1078"
df[df$improve_chem_id == drg & df$improve_sample_id %in% same_samps_1078,] %>% 
    ggplot(aes(x = auc, fill = source)) + 
    geom_histogram(alpha = 0.6, position = "identity", bins = 35) + 
    facet_wrap(~source) + 
    labs(
        title = glue("{drg} in Different Studies with Same Samples"), 
    )

```


```{r}
model <- aov(auc~source, data=data_1078)

summary(model)
```

```{r}
TukeyHSD(model)
```

## Looking at 988 same samples auc distribution

```{r}
df[df$improve_chem_id == "Drug_988",] %>% 
    group_by(source) %>% 
    summarize(
        num_samples = n_distinct(improve_sample_id)
    )
    
```

```{r}
d988_GDSCv1 <- df[df$improve_chem_id == "Drug_988" & df$source == "GDSCv1",]
d988_CTRPv2 <- df[df$improve_chem_id == "Drug_988" & df$source == "CTRPv2",]
d988_GDSCv2 <- df[df$improve_chem_id == "Drug_988" & df$source == "GDSCv2",]
d988_gCSI <- df[df$improve_chem_id == "Drug_988" & df$source == "gCSI",]
```

```{r}
same_samp_988 <- intersect(d988_GDSCv1$improve_sample_id, intersect(d988_CTRPv2$improve_sample_id, intersect(d988_GDSCv2$improve_sample_id, d988_gCSI$improve_sample_id)))
```

## All samps
```{r}
drg = "Drug_988"
df[df$improve_chem_id == drg,] %>% 
    ggplot(aes(x = auc, fill = source)) + 
    geom_histogram(alpha = 0.6, position = "identity", bins = 35) + 
    facet_wrap(~source) + 
    labs(
        title = glue("{drg} in Different Studies with ALL Samples"), 
    )
```

# Same samps

```{r}
drg = "Drug_988"
df[df$improve_chem_id == drg & df$improve_sample_id %in% same_samp_988,] %>% 
    ggplot(aes(x = auc, fill = source)) + 
    geom_histogram(alpha = 0.6, position = "identity", bins = 35) + 
    facet_wrap(~source) + 
    labs(
        title = glue("{drg} in Different Studies with Same Samples"), 
    )

# ggsave("drug988_by_study_samesamps.png", path = "/Users/jakegwinn/Documents/argonne_all/argonne/plots")
```

## Looking at difference between studies with sample measurments

```{r}
drg = "Drug_988"
d_988 <- df[df$improve_chem_id == drg & df$improve_sample_id %in% same_samp_988,]
d_full_988 <- df[df$improve_chem_id == drg,]
```

```{r}
d_988 %>% 
    group_by(improve_sample_id) %>% 
    summarize(
        num = n()
    ) %>% 
    arrange(desc(num))
```


```{r}
samp_988 = "ACH-000927"

d_988[d_988$improve_sample_id == samp_988,] %>% 
    dplyr::select(source, improve_chem_id, improve_sample_id, auc) %>% 
    arrange(source, desc(auc))
    

d_988 %>% 
    dplyr::select(source, improve_sample_id, auc) %>% 
    group_by(improve_sample_id) %>% 
    summarize(
        low = min(auc), 
        high = max(auc), 
        range = high - low
    ) %>% 
    arrange(desc(range))
    
```

```{r}
d_988 %>% 
    group_by(source) %>% 
    summarize(count = n())
```

```{r}
model_988 <- aov(auc~source, data=d_988)

summary(model_988)
```

```{r}
TukeyHSD(model_988)
```

```{r}
# plot(TukeyHSD(model_988), las = 2)
```

```{r}
df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        num_studies = n_distinct(source), 
        num_obs = n(), 
        var_auc = var(auc)
    ) %>% 
    arrange(desc(var_auc))
```

```{r}
by_study_box("Drug_988")
```

```{r}
df %>% 
    group_by(source) %>% 
    summarize(
        count = n()
    ) %>% 
    arrange(count) %>% 
    ggplot(aes(x = source, y = count)) + 
    geom_bar(stat="identity", position = "dodge", ) + 
    labs(
        title = "Frequency of Study Observations"
    )
```

# Study Analysis

```{r}
df %>% 
    group_by(source) %>% 
    summarize(
        num_obs = n()
    )
```


```{r}
df %>% 
    ggplot(aes(x = auc)) + 
    geom_histogram(bins = 70) + 
    facet_wrap(~source)
```

```{r}
ccle_df <- df[df$source == "CCLE", ]
ctrp_df <- df[df$source == "CTRPv2", ]
gcsi_df <- df[df$source == "gCSI", ]

gdsc1_df <- df[df$source == "GDSCv1", ]
gdsc2_df <- df[df$source == "GDSCv2", ]

```

```{r}
ccle_df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs))

ctrp_df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs))

gcsi_df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs))

gdsc1_df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs))

gdsc2_df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs))

```

```{r}
df %>% 
    group_by(improve_chem_id) %>% 
    summarize(
        obs = n()
    ) %>% 
    arrange(desc(obs)) 
```

```{r}
df %>% 
    dplyr::select(source) %>% 
    unique()
```

